### 2004-07-09 Documentation/part1.tex
*** empty log message ***

\chapter{A \GLOBES\ tour}
\label{chapter:tour}

@@

In this first chapter of the user's manual, we assume that the \GLOBES\ software is readily installed on your computer system. For the installation,
see \App~\ref{app:installation} and the {\tt INSTALL} file in the
software package. We demonstrate how to load pre-defined experiments 
and introduce the basic concepts of \GLOBES . We do not go
into details of the programming language, which means that standard parts
of the program code common to all of the examples in the following chapters are, in general, omitted.
An example of a minimal \GLOBES\ program in C can be found on page~\pageref{ex:c}. Furthermore, the files of the examples in this part can be found in the {\tt Example} subdirectory of your \GLOBES\ distribution. \index{Examples} Since it depends on your installed
configuration, we refer to the {\tt README} file in the \GLOBES\ main directory, and the comments in the Makefile of the {\tt Example} subdirectory for how to complile the example files. Make sure that
the data files (\AEDL\ and supporting files) in the {\tt data} 
subdirectory can be found by the
examples. In case of doubt, you can simply copy the necessary files
into the working directory where the examples are executed.

@@

\example{Using \GLOBES\ with C}{\label{ex:c}
\index{C-Code}
Here comes the C-code sceleton, which is
(more or less) common to all of our \GLOBES\ examples:
\begin{quote}
{\tt {\footnotesize
\#include <stdio.h> \\
\#include <stdlib.h> \\
\#include <math.h> \\
\#include <string.h> \\
\\
\#include <globes/globes.h> \hspace*{0.5cm} /* Include GLoBES library */ \\
\\
\#include "myio.h" \hspace*{0.5cm} /* Include "housemade" I/O-routines */ \\
\\
/* If filename given, write to file; if empty, to screen: */ \\
char MYFILE[]="testX.dat"; \\
\\
int main(int argc, char *argv[]) \\
\{  \\
\\
 \hspace*{0.5cm} glbInit(argv[0]); \hspace*{0.5cm} /* Initialize GLoBES library */ \\
\\  
  \hspace*{0.5cm} glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\
  \hspace*{1cm} \&glb\_num\_of\_exps); \hspace*{0.5cm} /* Initialize experiment NuFact.glb */\\
\\  
  \hspace*{0.5cm}  /* Initialize housemade output function */\\
  \hspace*{0.5cm} 
   InitOutput(MYFILE,"Format: ... ... ... $\backslash$n"); \\
\\  
  \hspace*{0.5cm} /* Initialize parameter vector(s) */ \\
  \hspace*{0.5cm} glb\_params true\_values = glbAllocParams(); \\
  \hspace*{0.5cm} /* ... */ \\
\\  
 \hspace*{0.5cm} /* Assign: theta12,theta13,theta23,deltacp,dm2solar,dm2atm */ \\
   \hspace*{0.5cm}  
     glbDefineParams(true\_values,\\
     \hspace*{1.5cm}asin(sqrt(0.8))/2,asin(sqrt(0.001))/2,M\_PI/4,M\_PI/2,7e-5,2e-3); \\

@@

\underline{$\beta$-Beams:} & \multicolumn{3}{l}{In preparation} \\[0.1cm]
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:experiments} \index{Experiment data files (table)} Pre-defined experiment prototypes, their filenames (to be used in \GLB{glbInitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
different implementation of the energy threshold function). Note that
some of these experiments are outdated in terms of integrated luminosities,
baseline, fluxes, efficiencies, or other factors.}
\end{table}

\section{Initialization of \GLOBES}

Before one can use \GLOBES , one has to initialize the \GLOBES\
library \GLB{libglobes}:
\begin{function}
\index{Initialization of {\tt libglobes}}
\GLBNS{glbInit} {\tt void glbInit(char *name)} initializes the library {\tt libglobes} and has
to be called in the beginning of each \GLOBES\ program. It takes the
name {\tt name} of the program as a string to initialize the error handling
functions. In many cases, it is sufficient to use the first
argument from the command line as the program name (such as in example on page~\pageref{ex:c}).
\end{function}

@@

For storing the experiments, \GLOBES\ uses the initially empty list of experiments \GLB{glb\_experiment\_list}. To add a pre-defined experiment to this list, one can use the function {\tt glbInitExperiment}:
\begin{function}
\index{Experiment initialization}
\GLBNS{glbInitExperiment}
{\tt int glbInitExperiment(char *inf, glb\_exp *in, int *counter)}
 adds a single experiment with the filename {\tt inf} to the list of currently loaded experiments. The {\tt counter} is a pointer to the 
 variable containg the number of experiments, and the experiment {\tt in}
 points to the beginning of the experiment list. The function returns
 zero if it was successful. 
\end{function}
Normally, a typical  call of {\tt glbInitExperiment} is 
\begin{quote}
{\tt glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0],\\  \hspace*{8cm} \&glb\_num\_of\_exps); }
\end{quote}
In this case, the experiment in the file {\tt NuFact.glb} is added to the internal global list of experiments, and the experiment counter is increased. The
experiment then has the number {\tt glb\_num\_of\_exps}-1. The elements
of the experiment list have the type \GLB{glb\_exp}, which the
user will not need to access directly in most cases. The experiment definition files, which usually end with {\tt .glb}, and any
supporting files, are first of all searched in the current directory, and then in the global path variable \GLB{GLB\_PATH} set in your environment.
A list of pre-defined experiment prototypes, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}. If the program cannot find these files, or some of them are syntactially not correct, it will break at this place.

One can also remove all experiments from the evaluation list at running
 time:
\begin{function}
\GLBNS{glbClearExperimentList}
{\tt void glbClearExperimentList()} removes all experiments from the internal list and resets all counters.   
\end{function}
Note that changing the number of experiments requires a new initialization
of all parameters of the types \GLB{glb\_params} and \GLB{glb\_projection}
if the number of experiments changes, since these parameter structures internally carry lists for the matter densities of all experiments. Similarly, once should never call {\tt glbAlloc...} before the
experiment initialization.

\begin{table}[tb]
\begin{center}
\begin{tabular}{llp{7cm}}
\hline
Quantities & Examples & Units \\
\hline
Angles & $\theta_{13}$, $\theta_{12}$, $\theta_{23}$, $\deltacp$ & Radians  \\
Mass squared differences & $\sdm$, $\ldm$ & $\mathrm{eV}^2$ \\
Matter densitities & $\rho_i$ & $\mathrm{g}/\mathrm{cm}^3$ \\
Baseline lengths & $L_i$ & $\mathrm{km}$ \\
Energies & $E_\nu$ & $\mathrm{GeV}$ \\  
Fiducial masses & $m_{\mathrm{Det}}$ & $\mathrm{t}$ (reactor exp.) or $\mathrm{kt}$ (accelerator exp.), \newline  depends on experiment definition \\
Time intervals & $t_{\mathrm{run}}$ & $\mathrm{yr}$ \\
Source powers & $P_{\mathrm{Source}}$ & Useful parent particle decays/$\mathrm{yr}$ \newline (Neutrino factory, $\beta$-Beam), \newline $\mathrm{GW}$ thermal power (reactor exps.), \newline or
$\mathrm{MW}$ target power (superbeams); \newline
depends on flux definition
 \\
% Integrated luminosities & $m_{\mathrm{Det}} \, t_{\mathrm{run}}$ & $\mathrm{kt \cdot yr}$ \\
Cross sections/E & $\sigma_{\mathrm{CC}}/E$ & $10^{-38} \, \mathrm{cm^2}/\mathrm{GeV^2}$ \\
\hline
\end{tabular}
\mycaption{\label{tab:units} \index{Units in \GLOBES } Quantities used in \GLOBES , examples of these quantities, and their standard units in the application software.}
\end{center}
\end{table}

\section{Units in \GLOBES\ and the integrated luminosity}
\label{sec:luminosity}

While interacting with the user interface of \GLOBES , parameters are transferred to and from the \GLOBES\ library. In \GLOBES , one set of units 
for each type of quantity is used in order to avoid confusion about the definition of individual parameters. \tabl{units} summarizes the units of the most important quantities. In principle, the event rates are
proportional to the product of source power $\times$ target mass $\times$
 running time, which we call ``integrated luminosity''. Since especially the
 definition of the source power depends on the experiment type, the quantities of the three luminosity components
 are not unique and depend on the experiment definition. Usually,
 one uses detector masses in kilotons for beam experiments,
 and detector masses in tons for reactor experiments. Running times
 are normally given in years, where it is often assumed that the 
 experiment runs $100\%$ of the year. Thus, for shorter running periods,
 the running times need to be renormalized. Source powers are
 usually useful parent particle decays per year (neutrino factories,
 $\beta$-beams), target power in megawatts (superbeams), or thermal
 reactor power in gigawatts (reactor experiments).
%
Since the pre-defined experiments in \tabl{experiments} are given for specfific luminosities, it is useful to read out and change these
 parameters of the individual experiments:
\begin{function}
\GLBNS{glbSetSourcePower}
{\tt void glbSetSourcePower(double power, int exp, int fluxno)}
 sets the source power of experiment number {\tt exp} and flux number
 {\tt fluxno} to {\tt power}. The definition of the source power depends on the experiment type as described above.
\end{function}
\begin{function}
\GLBNS{glbGetSourcePower}
{\tt double glbGetSourcePower(int exp, int fluxno)}
 returns the source power of experiment number {\tt exp} and flux number
 {\tt fluxno}. 
\end{function}
\begin{function}
\GLBNS{glbSetRunningTime}
{\tt void glbSetRunningTime(double time, int exp, int fluxno)}
 sets the running time of experiment number {\tt exp} and flux number
 {\tt fluxno} to {\tt time} years.
\end{function}
\begin{function}
\GLBNS{glbGetRunningTime}
{\tt double glbGetRunningTime(int exp, int fluxno)}
 returns the running time of experiment number {\tt exp} and flux number
 {\tt fluxno}.
\end{function}
\begin{function}
 \GLBNS{glbSetTargetMass}
{\tt void glbSetTargetMass(double mass,int exp)}
sets the fiducial detector mass of experiment number {\tt exp} to {\tt mass} tons or kilotons (depending on the experiment definition).
\end{function}
\begin{function}
 \GLBNS{glbGetTargetMass}
{\tt double glbGetTargetMass(int exp)}
returns the fiducial detector mass of experiment number {\tt exp}.
\end{function}
Thus, these functions also demonstrate how to use the assigned experiment number and others. These numbers run from $0$ to the number of experiments-1, fluxes-1, \etc, where the individual elements are numbered in the order of their appearance. 
 Note that the source power and running time are quantities defined
together with the neutrino flux, whereas the target mass scales the whole
experiment.

\section{Handling oscillation parameter vectors}

Before we can set the simulated event rates or access any oscillation
parameters, we need to become familiar with the concept \GLOBES\ uses
for oscillation parameters. In order to transfer sets of oscillation
parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$, the parameter type \GLB{glb\_params\_type} is used. In general, pointers to this structure are of the type \GLB{glb\_params}, and are often
transferred to and from \GLOBES\ functions. Therefore, the memory space
for these vectors has to be reserved before they can be used, and it has
to be returned afterwards. \GLOBES\ functions usually use the pointers
of the type {\tt glb\_params} for the input or output to the functions. As
an input parameter, the pointer has to be created and point towards a valid
parameter structure, where the oscillation parameters are read from. As
an output parameter, the pointer has to be created, too, and point towards
a structure which will contain the return values will be written to.
This parameter transfer concept seems to be very sophisticated, but, as we
will see in the next chapters, it hides a lot of complicated parameter 
mappings which otherwise need to be done by the user. For example, not
only the oscillation parameters are stored in the pointer structure, but also information on the matter
densities of all of the initialized experiments. Since \GLOBES\ treats
the matter density as a free parameter known with some external precision
to include matter density uncertainties,
the minimizers also use fit values and external errors for the matter
densities of all loaded experiments. More precisely, the matter density
profile of each experiment $i$ is multiplied by a scaling factor
$\hat{\rho_i}$, \index{Matter density scaling factor} 
which is stored in the density information of {\tt glb\_params}. 
Each of
these scaling factors has $1.0$ as pre-defined value. Since it
is in most cases not necessary to change this value, the user does
not need to take care of it. For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$. For a matter density profile, it acts as an overall normalization factor: The matter density in each layer is multiplied by this factor. In most cases one wants to take a scaling factor of $1.0$ here, which simply means taking the matter density profile as it is given in the experiment definition. For the treatment of correlations, however,
an external precision of the scaling factor might be used to include the
correlations with the matter density uncertainty. For more details, we
refer to the following sections.
Note that the {\tt glb\_params}
structures must not be initialized before all experiments are loaded,
since the number of matter densities can only be determined after the
experiments are initialized. 
Similarly, any change in the number of experiments requires that the
parameter structures be re-initialized, \ie, destroyed and created again.

@@

Similarly, the density parameters or iterations can be accessed:
\begin{function}
\GLBNS{glbSetDensityParams}
{\tt glb\_params glbSetDensityParams(glb\_params in, double dens, int which)} sets the density parameter {\tt which} in the structure {\tt in}
to the value {\tt dens}. If the assignment was unsuccessful, the 
function returns {\tt NULL}. If {\tt GLB\_ALL} is used for {\tt which},
the density parameters of all experiments will be set accordingly.
\end{function}
\begin{function}
\GLBNS{glbGetDensityParams}
{\tt double glbGetDensityParams(glb\_params in, int which)}
returns the value of the density parameter {\tt which} in the 
structure {\tt in}.
\end{function}
\begin{function}
\GLBNS{glbSetIteration}
{\tt glb\_params glbSetIteration(glb\_params in, int iter)} sets the number of iterations in the structure {\tt in}
to the value {\tt iter}. If the assignment was unsuccessful, the 
function returns {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{glbGetIteration}
{\tt int glbGetIteration(glb\_params in)}
returns the value of the number of iterations in the 
structure {\tt in}.
\end{function}
In total, the parameter vector handling in a program normally has the
following order:
\begin{quote}
{\tt
glbInitExperiment(...); \\
/* ... more initializations ... */ \\
\\
glb\_params vector1 = glbAllocParams(); \\
/* ... more vectors allocated ... */ \\
\\
/* Program code: assign and use vectors */ \\
\\
glbFreeParams(vector1); \\
/* ... more vectors destroyed ... */ \\
\\
/* ... end of program or glbClearExperimentList ... */ 
}
\end{quote}

\section{Computing the simulated data}

@@

In order to keep track of the used version of \GLOBES , the software
provides a number of functions to check the \GLOBES\ and experiment
versions. It is up to the user to implement mechanisms into the
program and \AEDL\ files to check wether
\begin{itemize}
\item
 The program should only run with this specific version of \GLOBES
\item
 The program can only run with a minimum version of \GLOBES
\item
 The program can only run up to a certain \GLOBES\ version.
\end{itemize}
The same holds for \AEDL\ files: For example, 
some features may not be supported by earlier versions of \GLOBES\ anymore.
The program can then check the version of the \AEDL\ file and break if it is
too old.

@@

\chapter[Calculating $\chi^2$ with systematics only]{Calculating $\boldsymbol{\chi^2}$ with systematics only}

\index{Systematics}
Calculating a $\chi^2$-value with or without systematics, but no correlations and degeneracies, is the simplest and fastest possibility to obtain high-level information on an experiment. In general, \GLOBES\ uses the six independent oscillation parameters $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$, as well as the matter density 
scaling factor $\hat{\rho}$ of each experiment.\index{Matter density
scaling factor} Thus, there are six plus the number of experiments parameters determining the rate vectors. Using the matter density scaling factors  in addition to the oscillation parameters will allow the simulation of the correlations with matter density uncertainties: In this
approach, the matter density profile normalization $\hat{\rho}$ can 
be treated
as parameter to be measured by the experiment, where an external precision
given by observations is imposed (typically up to $5\%$). 
 For this section, it is important to keep in mind
that there are more parameters than just the oscillation parameters
determining the simple $\chi^2$. However, as we have described in the
last section, the mechanism for the matter density scaling factors
is hidden in the defintion of \GLB{glb\_params}: Each of the scaling
factors is initially set to $1.0$. Therefore, for the calculation of
$\chi^2$ with systematics only, we do not have to care about the
matter density scaling factors.

@@

Keeping all oscillation parameters and matter density scaling factors fixed,
 one can use the following functions to obtain the total $\chi^2$ of all 
 specified oscillation channels including systematics:
\begin{function} 
\GLBNS{glbChiSys}
{\tt double glbChiSys(const glb\_params in,int exp, int rule)} returns
the $\chi^2$ for the (fixed) oscillation parameters {\tt in}, the
experiment number {\tt exp}, and the rule number {\tt rule}. For all
experiments or rules, use \GLB{GLB\_ALL} as parameter value.
\end{function}
Note that the result of {\tt glbChiSys} for all experiments or rules
corresponds to the sum of all of the individual {\tt glbChiSys} calls. 
This equality will not hold for the minimizors in the next sections anymore. 
 An example how to use  {\tt glbChiSys} can be found on page~\pageref{ex:corrth13dcp}.

>>>>

\chapter{A \GLOBES\ tour}
\label{chapter:tour}
\index{GLoBES tour}

@@

\index{Installation}
In this first chapter of the user's manual, we assume that the \GLOBES\ software is readily installed on your computer system. For the installation,
see \App~\ref{app:installation} and the {\tt INSTALL} file in the
software package. We demonstrate how to load pre-defined experiments 
and introduce the basic concepts of \GLOBES . We do not go
into details of the programming language, which means that standard parts
of the program code common to all of the examples in the following chapters are, in general, omitted.
An example of a minimal \GLOBES\ program in C can be found on page~\pageref{ex:c}. Furthermore, the files of the examples in this part can be found in the {\tt Example} subdirectory of your \GLOBES\ distribution. \index{Examples} Since it depends on your installed
configuration, we refer to the {\tt README} file in the \GLOBES\ main directory, and the comments in the Makefile of the {\tt Example} subdirectory for how to compile the example files. Make sure that
the data files (\AEDL\ and supporting files) in the {\tt data} 
subdirectory can be found by the
examples. In case of doubt, you can simply copy the necessary files
into the working directory where the examples are executed.

@@

\example{Using \GLOBES\ with C}{\label{ex:c}
\index{C-Code}
\index{Program}
Here comes the C-code skeleton, which is
(more or less) common to all of our \GLOBES\ examples:
\begin{quote}
{\tt {\footnotesize
\#include <stdio.h> \\
\#include <stdlib.h> \\
\#include <math.h> \\
\#include <string.h> \\
\\
\#include <globes/globes.h> \hspace*{0.5cm} /* Include GLoBES library */ \\
\\
\#include "myio.h" \hspace*{0.5cm} /* Include "housemade" I/O-routines */ \\
\\
/* If filename given, write to file; if empty, to screen: */ \\
char MYFILE[]="testX.dat"; \\
\\
int main(int argc, char *argv[]) \\
\{  \\
\\
 \hspace*{0.5cm} glbInit(argv[0]); \hspace*{0.5cm} /* Initialize GLoBES library */ \\
\\  
  \hspace*{0.5cm} glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\
  \hspace*{1cm} \&glb\_num\_of\_exps); \hspace*{0.5cm} /* Initialize experiment NuFact.glb */\\
\\  
  \hspace*{0.5cm}  /* Initialize housemade output function */\\
  \hspace*{0.5cm} 
   InitOutput(MYFILE,"Format: ... ... ... $\backslash$n"); \\
\\  
  \hspace*{0.5cm} /* Initialize parameter vector(s) */ \\
  \hspace*{0.5cm} glb\_params true\_values = glbAllocParams(); \\
  \hspace*{0.5cm} /* ... */ \\
\\  
 \hspace*{0.5cm} /* Assign: theta12,theta13,theta23,deltacp,dm2solar,dm2atm */ \\
   \hspace*{0.5cm}  
     glbDefineParams(true\_values,\\
     \hspace*{1.5cm}asin(sqrt(0.8))/2,asin(sqrt(0.001))/2,M\_PI/4,M\_PI/2,7e-5,2e-3); \\

@@

\underline{$\beta$-Beams:} & \multicolumn{3}{l}{In preparation} \\[0.1cm]
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:experiments} \index{Experiment files (table)}
\index{{\tt glb}-files} 
Pre-defined experiment prototypes, their filenames (to be used in \GLB{glbInitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
different implementation of the energy threshold function). Note that
some of these experiments are outdated in terms of integrated luminosities,
baseline, fluxes, efficiencies, or other factors.}
\end{table}

\section{Initialization of \GLOBES}
\index{Initialization}

Before one can use \GLOBES , one has to initialize the \GLOBES\
library \GLB{libglobes}: \index{Initialization: \GLOBES\ library}
\begin{function}
\index{Initialization of {\tt libglobes}}
\GLBNS{glbInit} {\tt void glbInit(char *name)} initializes the library {\tt libglobes} and has
to be called in the beginning of each \GLOBES\ program. It takes the
name {\tt name} of the program as a string to initialize the error handling
functions. In many cases, it is sufficient to use the first
argument from the command line as the program name (such as in example on page~\pageref{ex:c}).
\end{function}

@@

\index{Experiment list}
\index{Initialization: Experiments}
For storing the experiments, \GLOBES\ uses the initially empty list of experiments \GLB{glb\_experiment\_list}. To add a pre-defined experiment to this list, one can use the function {\tt glbInitExperiment}:
\begin{function}
\index{Experiment initialization}
\GLBNS{glbInitExperiment}
{\tt int glbInitExperiment(char *inf, glb\_exp *in, int *counter)}
 adds a single experiment with the filename {\tt inf} to the list of currently loaded experiments. The {\tt counter} is a pointer to the 
 variable containing the number of experiments, and the experiment {\tt in}
 points to the beginning of the experiment list. The function returns
 zero if it was successful. 
\end{function}
Normally, a typical  call of {\tt glbInitExperiment} is 
\begin{quote}
{\tt glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0],\\  \hspace*{8cm} \&glb\_num\_of\_exps); }
\end{quote}
In this case, the experiment in the file {\tt NuFact.glb} is added to the internal global list of experiments, and the experiment counter is increased. The
experiment then has the number {\tt glb\_num\_of\_exps}-1. The elements
of the experiment list have the type \GLB{glb\_exp}, which the
user will not need to access directly in most cases. The experiment definition files, which usually end with {\tt .glb}, and any
supporting files, are first of all searched in the current directory, and then in the global path variable \GLB{GLB\_PATH} set in your environment.
A list of pre-defined experiment prototypes, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}. If the program cannot find these files, or some of them are syntactically not correct, it will break at this place.

\index{Clear experiment list}
\index{Delete experiments}
One can also remove all experiments from the evaluation list at running
 time:
\begin{function}
\GLBNS{glbClearExperimentList}
{\tt void glbClearExperimentList()} removes all experiments from the internal list and resets all counters.   
\end{function}
Note that changing the number of experiments requires a new initialization
of all parameters of the types \GLB{glb\_params} and \GLB{glb\_projection}
if the number of experiments changes, since these parameter structures internally carry lists for the matter densities of all experiments. Similarly, once should never call {\tt glbAlloc...} before the
experiment initialization.

\begin{table}[tb]
\begin{center}
\begin{tabular}{llp{7cm}}
\hline
Quantities & Examples & Units \\
\hline
Angles & $\theta_{13}$, $\theta_{12}$, $\theta_{23}$, $\deltacp$ & Radians  \\
Mass squared differences & $\sdm$, $\ldm$ & $\mathrm{eV}^2$ \\
Matter densities & $\rho_i$ & $\mathrm{g}/\mathrm{cm}^3$ \\
Baseline lengths & $L_i$ & $\mathrm{km}$ \\
Energies & $E_\nu$ & $\mathrm{GeV}$ \\  
Fiducial masses & $m_{\mathrm{Det}}$ & $\mathrm{t}$ (reactor exp.) or $\mathrm{kt}$ (accelerator exp.), \newline  depends on experiment definition \\
Time intervals & $t_{\mathrm{run}}$ & $\mathrm{yr}$ \\
Source powers & $P_{\mathrm{Source}}$ & Useful parent particle decays/$\mathrm{yr}$ \newline (Neutrino factory, $\beta$-Beam), \newline $\mathrm{GW}$ thermal power (reactor exps.), \newline or
$\mathrm{MW}$ target power (superbeams); \newline
depends on flux definition
 \\
% Integrated luminosities & $m_{\mathrm{Det}} \, t_{\mathrm{run}}$ & $\mathrm{kt \cdot yr}$ \\
Cross sections/E & $\sigma_{\mathrm{CC}}/E$ & $10^{-38} \, \mathrm{cm^2}/\mathrm{GeV^2}$ \\
\hline
\end{tabular}
\mycaption{\label{tab:units} \index{Units in \GLOBES\ (table)} Quantities used in \GLOBES , examples of these quantities, and their standard units in the application software.}
\end{center}
\end{table}

\section{Units in \GLOBES\ and the integrated luminosity}
\label{sec:luminosity}
\index{Integrated luminosity}
\index{Detector mass} \index{Source power} \index{Running time}

While interacting with the user interface of \GLOBES , parameters are transferred to and from the \GLOBES\ library. In \GLOBES , one set of units 
for each type of quantity is used in order to avoid confusion about the definition of individual parameters. \tabl{units} summarizes the units of the most important quantities. In principle, the event rates are
proportional to the product of source power $\times$ target mass $\times$
 running time, which we call ``integrated luminosity''. Since especially the
 definition of the source power depends on the experiment type, the quantities of the three luminosity components
 are not unique and depend on the experiment definition. Usually,
 one uses detector masses in kilotons for beam experiments,
 and detector masses in tons for reactor experiments. Running times
 are normally given in years, where it is often assumed that the 
 experiment runs $100\%$ of the year. Thus, for shorter running periods,
 the running times need to be renormalized. Source powers are
 usually useful parent particle decays per year (neutrino factories,
 $\beta$-beams), target power in mega watts (superbeams), or thermal
 reactor power in giga watts (reactor experiments).
%
Since the pre-defined experiments in \tabl{experiments} are given for specific luminosities, it is useful to read out and change these
 parameters of the individual experiments:
\begin{function}
\GLBNS{glbSetSourcePower}
{\tt void glbSetSourcePower(double power, int exp, int fluxno)}
 sets the source power of experiment number {\tt exp} and flux number
 {\tt fluxno} to {\tt power}. The definition of the source power depends on the experiment type as described above.
\end{function}
\begin{function}
\GLBNS{glbGetSourcePower}
{\tt double glbGetSourcePower(int exp, int fluxno)}
 returns the source power of experiment number {\tt exp} and flux number
 {\tt fluxno}. 
\end{function}
\begin{function}
\GLBNS{glbSetRunningTime}
{\tt void glbSetRunningTime(double time, int exp, int fluxno)}
 sets the running time of experiment number {\tt exp} and flux number
 {\tt fluxno} to {\tt time} years.
\end{function}
\begin{function}
\GLBNS{glbGetRunningTime}
{\tt double glbGetRunningTime(int exp, int fluxno)}
 returns the running time of experiment number {\tt exp} and flux number
 {\tt fluxno}.
\end{function}
\begin{function}
 \GLBNS{glbSetTargetMass}
{\tt void glbSetTargetMass(double mass,int exp)}
sets the fiducial detector mass of experiment number {\tt exp} to {\tt mass} tons or kilotons (depending on the experiment definition).
\end{function}
\begin{function}
 \GLBNS{glbGetTargetMass}
{\tt double glbGetTargetMass(int exp)}
returns the fiducial detector mass of experiment number {\tt exp}.
\end{function}
Thus, these functions also demonstrate how to use the assigned experiment number and others. These numbers run from $0$ to the number of experiments-1, fluxes-1, \etc, where the individual elements are numbered in the order of their appearance. 
 Note that the source power and running time are quantities defined
together with the neutrino flux, whereas the target mass scales the whole
experiment.

\section{Handling oscillation parameter vectors}
\index{Oscillation parameter vectors}

Before we can set the simulated event rates or access any oscillation
parameters, we need to become familiar with the concept \GLOBES\ uses
for oscillation parameters. In order to transfer sets of oscillation
parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$, the parameter type \GLB{glb\_params\_type} is used. In general, pointers to this structure are of the type \GLB{glb\_params}, and are often
transferred to and from \GLOBES\ functions. Therefore, the memory space
for these vectors has to be reserved before they can be used, and it has
to be returned afterwards. \GLOBES\ functions usually use the pointers
of the type {\tt glb\_params} for the input or output to the functions. As
an input parameter, the pointer has to be created and point towards a valid
parameter structure, where the oscillation parameters are read from. As
an output parameter, the pointer has to be created, too, and point towards
a structure which will contain the return values will be written to.
This parameter transfer concept seems to be very sophisticated, but, as we
will see in the next chapters, it hides a lot of complicated parameter 
mappings which otherwise need to be done by the user. For example, not
only the oscillation parameters are stored in the pointer structure, but also information on the matter
densities of all of the initialized experiments. Since \GLOBES\ treats
the matter density as a free parameter known with some external precision
to include matter density uncertainties,
the minimizers also use fit values and external errors for the matter
densities of all loaded experiments. More precisely, the matter density
profile of each experiment $i$ is multiplied by a scaling factor
$\hat{\rho_i}$, \index{Matter density scaling factor}
\index{Matter density profile} 
which is stored in the density information of {\tt glb\_params}. 
Each of
these scaling factors has $1.0$ as pre-defined value. Since it
is in most cases not necessary to change this value, the user does
not need to take care of it. For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$. For a matter density profile, it acts as an overall normalization factor: The matter density in each layer is multiplied by this factor. In most cases one wants to take a scaling factor of $1.0$ here, which simply means taking the matter density profile as it is given in the experiment definition. For the treatment of correlations, however,
an external precision of the scaling factor might be used to include the
correlations with the matter density uncertainty. For more details, we
refer to the following sections.
Note that the {\tt glb\_params}
structures must not be initialized before all experiments are loaded,
since the number of matter densities can only be determined after the
experiments are initialized. 
Similarly, any change in the number of experiments requires that the
parameter structures be re-initialized, \ie, destroyed and created again.

@@

Similarly, the density parameters or iterations can be accessed:
\begin{function}
\GLBNS{glbSetDensityParams}
{\tt glb\_params glbSetDensityParams(glb\_params in, double dens, int which)} sets the density parameter {\tt which} in the structure {\tt in}
to the value {\tt dens}. If the assignment was unsuccessful, the 
function returns {\tt NULL}. If {\tt GLB\_ALL} is used for {\tt which},
the density parameters of all experiments will be set accordingly.
\end{function}
\begin{function}
\GLBNS{glbGetDensityParams}
{\tt double glbGetDensityParams(glb\_params in, int which)}
returns the value of the density parameter {\tt which} in the 
structure {\tt in}.
\end{function} 
\begin{function}
\GLBNS{glbSetIteration} \index{Iterations}
{\tt glb\_params glbSetIteration(glb\_params in, int iter)} sets the number of iterations in the structure {\tt in}
to the value {\tt iter}. If the assignment was unsuccessful, the 
function returns {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{glbGetIteration}
{\tt int glbGetIteration(glb\_params in)}
returns the value of the number of iterations in the 
structure {\tt in}.
\end{function}
In total, the parameter vector handling in a program normally has the
following order:\index{Parameter vector handling}
\begin{quote}
{\tt
glbInitExperiment(...); \\
/* ... more initializations ... */ \\
\\
glb\_params vector1 = glbAllocParams(); \\
/* ... more vectors allocated ... */ \\
\\
/* Program code: assign and use vectors */ \\
\\
glbFreeParams(vector1); \\
/* ... more vectors destroyed ... */ \\
\\
/* ... end of program or glbClearExperimentList ... */ 
}
\end{quote}

\section{Computing the simulated data}
\index{True values} \index{Simulated data} \index{Reference rate vector}
\index{Set oscillation parameters}

@@

In order to keep track of the used version of \GLOBES , the software
provides a number of functions to check the \GLOBES\ and experiment
versions. It is up to the user to implement mechanisms into the
program and \AEDL\ files to check whether
\begin{itemize}
\item
 The program should only run with this specific version of \GLOBES
\item
 The program can only run with a minimum version of \GLOBES
\item
 The program can only run up to a certain \GLOBES\ version.
\end{itemize}
The same holds for \AEDL\ files: For example, 
some features may not be supported by earlier versions of \GLOBES\ anymore.
The program can then check the version of the \AEDL\ file and break if it is
too old.

@@

\chapter[Calculating $\chi^2$ with systematics only]{Calculating $\boldsymbol{\chi^2}$ with systematics only}
\index{Systematics $\chi^2$}

Calculating a $\chi^2$-value with or without systematics, but no correlations and degeneracies, is the simplest and fastest possibility to obtain high-level information on an experiment. In general, \GLOBES\ uses the six independent oscillation parameters $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$, as well as the matter density 
scaling factor $\hat{\rho}$ of each experiment.\index{Matter density
scaling factor} Thus, there are six plus the number of experiments parameters determining the rate vectors. Using the matter density scaling factors  in addition to the oscillation parameters will allow the simulation of the correlations with matter density uncertainties: In this
approach, the matter density profile normalization $\hat{\rho}$ can 
be treated
as parameter to be measured by the experiment, where an external precision
given by observations is imposed (typically up to $5\%$). 
\index{Matter density uncertainty}
 For this section, it is important to keep in mind
that there are more parameters than just the oscillation parameters
determining the simple $\chi^2$. However, as we have described in the
last section, the mechanism for the matter density scaling factors
is hidden in the definition of \GLB{glb\_params}: Each of the scaling
factors is initially set to $1.0$. Therefore, for the calculation of
$\chi^2$ with systematics only, we do not have to care about the
matter density scaling factors.

@@

Keeping all oscillation parameters and matter density scaling factors fixed,
 one can use the following functions to obtain the total $\chi^2$ of all 
 specified oscillation channels including systematics:
\begin{function} 
\GLBNS{glbChiSys}
{\tt double glbChiSys(const glb\_params in,int exp, int rule)} returns
the $\chi^2$ for the (fixed) oscillation parameters {\tt in}, the
experiment number {\tt exp}, and the rule number {\tt rule}. For all
experiments or rules, use \GLB{GLB\_ALL} as parameter value.
\end{function}
Note that the result of {\tt glbChiSys} for all experiments or rules
corresponds to the sum of all of the individual {\tt glbChiSys} calls. 
This equality will not hold for the minimizers in the next sections anymore. 
 An example how to use  {\tt glbChiSys} can be found on page~\pageref{ex:corrth13dcp}.
